---
title: "ggplot2"
author: "D Li"
date: "November 10, 2015"
output: html_document
---

```{r, setup, include=FALSE, cache=FALSE}
library(knitr)
library(ggplot2)
opts_chunk$set(fig.align='center', cache=TRUE, tidy=TRUE, eval=FALSE, out.width='0.6\\textwidth') 
options(width = 85, replace.assign = TRUE)
```


Getting started with qplot (quick-plot)
=======================================

Basic use
---------

```{r}
library("ggplot2")
library(mgcv)
set.seed(1410)
dsmall <- diamonds[sample(nrow(diamonds), 100),]
## Basic use
qplot(x,y,data=,color=,shape=,size=,alpha=,geom=,method=,formula=,facets=,xlin=,ylim=,xlab=,ylab=,main=,sub=)
#color,shape,size,fill: associates the levels of variable with symbol color, shape, or size.
#method: "smooth","lm","gam" for generalized additive models(mgcv package), "rlm" for robust regression (MASS package).
#formula=y~x. use letters x and y, not the names of variables.
qplot(carat, price,data=diamonds) 
qplot(log(carat), log(price),data=diamonds) 
qplot(carat, x*y*z, data=diamonds)
#aesthetic attributes 
qplot(carat, price,data=diamonds, alpha=I(1/10)) 
#alpha aesthetics using I(), value from 0-comp transparent to 1, e.g.1/10.
#colour, shape 
qplot(carat,price,data=dsmall,colour=color) 
qplot(carat,price,data=dsmall,shape=cut)
```

Plot geom
---------

Geom, short for geometric object, describes the type of object that is
used to display the data. Here well introduce the most common and useful
geoms, organised by the dimensionality of data that they work with.

The following geoms enable you to investigate two-dimensional
relationships:

-   geom = "point" draws points to produce a scatterplot. This is
    the default when you supply both x and y arguments to qplot().

-   geom = "smooth" fits a smoother to the data and displays the
    smooth and its standard error.

-   geom = "boxplot" produces a box and whisker plot to summarise
    the distribution of a set of points.

-   geom = "path" and geom = "line" draw lines between the data
    points. Traditionally these are used to explore relationships
    between time and another variable, but lines may be used to join
    observations connected in some other way. A line plot is constrained
    to produce lines that travel from left to right, while paths can go
    in any direction.

For 1d distributions, your choice of geoms is guided by the variable
type:

-   For continuous variables, geom = "histogram" draws a histogram,
    geom = "freqpoly" a frequency polygon, and geom = "density"
    creates a density plot. The **histogram geom is the default when you
    only supply an x value to qplot()**.

-   For discrete variables, geom = "bar" makes a bar chart.

```{r}
#points and smooth
qplot(carat, price, data=dsmall,geom=c("point","smooth"))
qplot(carat, price, data=dsmall,geom=c("point","smooth"),se=FALSE) # turn off CI
qplot(carat, price, data=dsmall,geom=c("point","smooth"),span=0.5) # wiggliness
```

Methods for smooth:

-   for small n, default is: method= "loess". Wiggliness of the line
    is controlled by span from 0 (exceedingly wiggly) to 1 (not so
    wiggly, smooth).

-   You could also load the` mgcv` library and use
    `method="gam",formula=y~s(x)` to fit a generalised additive model.
    This is similar to using a spline with lm, but the degree of
    smoothness is estimated from the data. For large data, use the
    formula `y~s(x,bs="cs")`. This is used by default when there are
    more than 1,000 points.

-   `method="lm"` fits a linear model. The default will fit a straight line to your data,
    or you can specify `formula=y~poly(x,2)` to specify a degree 2
    polynomial, or better, load splines package and use a natural
    spline: `formula=y~ns(x,2)`.

-   `method = "rlm"` works like `lm`, but uses a robust fitting algorithm so that outliers dont
    affect the fit as much.

```{r}
library(mgcv)
qplot(carat, price, data=dsmall,geom=c("point","smooth"),method="gam", formula=y~s(x, bs="cs"))
```


How the values of the continuous variables vary with the levels of the
categorical variable? We can use boxplots and jittered points. For
jittered points, qplot offers the same control over aesthetics as it
does for a normal scatterplot: size, colour and shape. For boxplots you
can control the outline colour, the internal fill colour and the size of
the lines.

```{r}
#jitter and boxplot
qplot(color, price/carat, data=diamonds, geom="jitter", alpha=I(1/10))
qplot(color, price/carat, data=diamonds, geom="boxplot")
qplot(color, price/carat, data=diamonds, geom=c("jitter","boxplot"),alpha=I(0.5),colour=color)
```


Histogram and density plots show the distribution of a single variable.
For the density plot, the adjust argument controls the degree of
smoothness (high values of adjust produce smoother plots). For the
histogram, the binwidth argument controls the amount of smoothing by
setting the bin size. (Break points can also be specified explicitly,
using the breaks argument.)

```{r}
#histogram and density
qplot(carat,data=diamonds,geom="histogram", binwidth=0.1,xlim=c(0,3),colour = I("darkgreen"), fill = I("white")) 
qplot(carat,data=diamonds,geom="histogram", fill=color) 
qplot(carat,data=diamonds,geom="density") 
qplot(carat,data=diamonds,geom="density",adjust=.1) 
qplot(carat,data=diamonds,geom="density", colour=color)
```


Bar charts. The discrete analogue of histogram is the bar chart, geom =
"bar". The bar geom counts the number of instances of each class so
that you dont need to tabulate your values beforehand, as with barchart
in base R. If the data has already been tabulated or if youd like to
tabulate class members in some other way, such as by summing up a
continuous variable, you can use the weight geom.

```{r}
#barchart
qplot(color,data=diamonds, geom="bar") # try fill=color
qplot(color,data=diamonds, geom="bar",weight=carat)+scale_y_continuous("carat")
```

Line and path plots are typically used for time series data. Line plots
join the points from left to right, while path plots join them in the
order that they appear in the dataset (a line plot is just a path plot
of the data sorted by x value). Line plots usually have time on the
x-axis, showing how a single variable has changed over time. Path plots
show how two variables have simultaneously changed over time, with time
encoded in the way that the points are joined together.

```{r}
#Line and path graph
qplot(date,unemploy/pop,data=economics,geom="line")
year <- function(x) as.POSIXlt(x)$year + 1900 # e.g.1967
qplot(unemploy / pop, uempmed, data = economics,geom = c("point", "path")) 
qplot(unemploy / pop, uempmed, data = economics,geom = "path", colour = year(date)) + scale_area()
```


Faceting
--------

We have already discussed using aesthetics (colour and shape) to compare
subgroups, drawing all groups on the same plot. Faceting takes an
alternative approach: It creates tables of graphics by splitting the
data into subsets and displaying the same graph for each subset in an
arrangement that facilitates comparison. The default faceting method in
`qplot()` creates plots arranged on a grid specified by a faceting formula
which looks like row var\~col var. To facet on only one of columns or
rows, use . as a place holder. For example, row var \~ . will create a
single column with multiple rows. The `..density..` syntax is new. The
y-axis of the histogram does not come from the original data, but from
the statistical transformation that counts the number of observations in
each bin. Using `..density..` tells ggplot2 to map the density to the
y-axis instead of the default use of count.

```{r}
#Facet
qplot(carat, data = diamonds, facets = color ~ ., geom = "histogram", binwidth = 0.1, xlim = c(0, 3)) 
qplot(carat, ..density.., data = diamonds, facets = color ~ .,geom = "histogram", binwidth = 0.1, xlim = c(0, 3))
```


Other options
-------------

These are a few other qplot options to control the graphics appearance.
These all have the same effect as their plot equivalents: `xlim, ylim,
log, main, xlab, ylab`, etc.

Mastering the grammer
=====================

Scatterplot
-----------

```{r}
#scatterplot
qplot(displ, hwy, data = mpg, colour = factor(cyl))
```


A scatterplot represents each observation as a `point ()`, positioned
according to the value of two variables. As well as a horizontal and
vertical position, each point also has *a size, a colour and a shape*.
These attributes are called **aesthetics**, and are the properties that
can be perceived on the graphic.

*Points, lines and bars* are all examples of geometric objects, or
**geoms**. Geoms determine the type of the plot.

We need to convert them from data units (e.g., litres, miles per gallon
and number of cylinders) to physical units (e.g., pixels and colours)
that the computer can display. This conversion process is called
**scaling** and performed by scales.

A final step determines how the two positions (x and y) are combined to
form the final location on the plot. This is done by the coordinate
system, or **coord**. In most cases this will be Cartesian coordinates,
but it might be polar coordinates, or a spherical projection used for a
map.

Finally, we need to render this data to create the graphical objects
that are displayed on the screen. To create a complete plot we need to
combine graphical objects from three sources: the *data*, represented by
the point geom; the *scales and coordinate system*, which generate axes
and legends so that we can read values from the graph; and *plot
annotations*, such as the background and plot title.

A more complex plot
-------------------

```{r}
qplot(displ, hwy, data = mpg, facets = . ~ year) + geom_smooth()
```


This plot adds three new components to the mix: facets, multiple layers
and statistics. The facets and layers expand the data structure
described above: each facet panel in each layer has its own dataset. You
can think of this as a 3d array: the panels of the facets form a 2d
grid, and the layers extend upwards in the 3rd dimension. In this case
the data in the layers is the same, but in general *we can plot
different datasets on different layers*.

The smooth layer is different to the point layer because it doesnt
display the raw data, but instead displays a statistical transformation
of the data. Specifically, the smooth layer fits a smooth line through
the middle of the data. This requires an additional step in the process
described above: after mapping the data to aesthetics, the data is
passed to a statistical transformation, or **stat**, which manipulates
the data in some useful way.

As well as adding an additional step to summarise the data, we also need
some extra steps when we get to the scales. This is because we now have
multiple datasets (for the different facets and layers) and we need to
make sure that the scales are the same across all of them. Scaling
actually occurs in three parts: transforming, training and mapping.
Scale transformation occurs before statistical transformation so that
statistics are computed on the scale-transformed data. This ensures that
a plot of log(x) vs. log(y) on linear scales looks the same as x vs. y
on log scales. There are many different transformations that can be
used, including taking square roots, logarithms and reciprocals.After
the statistics are computed, each scale is trained on every dataset from
all the layers and facets. The training operation combines the ranges of
the individual datasets to get the range of the complete data. Without
this step, scales could only make sense locally and we wouldnt be able
to overlay different layers because their positions wouldnt line up.
Finally the scales map the data values into aesthetic values.

Faceting, a general case of the conditioned or trellised plots, makes it
easy to create small multiples each showing a different subset of the
whole dataset.

Data structures
---------------

A plot object is a list with components *data, mapping* (the default
aesthetic mappings), *layers, scales, coordinates* and *facet*. The plot
object has one other component we havent discussed yet: *options*.

Plots can be created in two ways: all at once with `qplot()`, as shown
in the previous chapter, or piece-by-piece with `ggplot()` and layer
functions. Once you have a plot object, there are a few things you can
do with it:

-   Render it on screen, with `print()`. This happens automatically when
    running interactively, but inside a loop or function, youll need to
    `print()` it yourself.

-   Render it to disk, with `ggsave()`.

-   Briefly describe its structure with `summary()`.

-   Save a cached copy of it to disk, with `save()`. This saves a
    complete copy of the plot object, so you can easily re-create that
    exact plot with `load()`. Note that data is stored inside the plot,
    so that if you change the data outside of the plot, and then redraw
    a saved plot, it will not be updated.

```{r}
p <- qplot(displ, hwy, data = mpg, colour = factor(cyl))
summary(p)
save(p, file = "plot.rdata") # Save plot object to disk
load("plot.rdata") # Load from disk
ggsave("plot.png", width = 5, height = 5) # Save png to disk
```


Build a plot layer by layer
===========================

Layering is the mechanism by which additional data elements are added to
a plot. Each layer can come from a different dataset and have a
different aesthetic mapping, allowing us to create plots that could not
be generated using `qplot()`, which permits only a single dataset and a
single set of aesthetic mappings.

Creating a plot
---------------

When we used `qplot()`, it did a lot of things for us: it created a plot
object, added layers, and displayed the result, using many default
values along the way. To create the plot object ourselves, we use
`ggplot()`. This has two arguments: **data** and aesthetic **mapping**.
These arguments set up defaults for the plot and can be omitted if you
specify data and aesthetics when adding each layer. You are already
familiar with aesthetic mappings from `qplot()`, and the syntax here is
quite similar, although you need to wrap the pairs of aesthetic
attribute and variable name in the `aes()` function.

```{r}
p <- ggplot(diamonds, aes(carat, price, colour = cut))
# This plot object cannot be displayed until we add a layer: there is
# nothing to see!
```


Layers
------

A minimal layer may do nothing more than specify a geom, a way of
visually representing the data.

```{r}
p <- p + layer(geom = "point")
```


This layer uses the plot defaults for data and aesthetic mapping and it
uses default values for two optional arguments: the statistical
transformation (the stat) and the position adjustment. A more fully
specified layer can take any or all of these arguments:
```{r}
# arguments for layers
layer(geom, geom_params, stat_params, data, mapping, position)
```


```{r, eval=TRUE, out.width='0.4 \ \textwidth',fig.show='hold'}
p <- ggplot(diamonds, aes(x = carat))
p1 <- p + layer(geom = "bar", geom_params = list(fill = "steelblue"), stat = "bin",
stat_params = list(binwidth = 2))
q <- ggplot(diamonds, aes(x = carat))
q <- p + layer(geom = "bar", geom_params = list(fill = "steelblue"), stat = "bin",
stat_params = list(binwidth = 0.5))
p1
q
```


This layer specification is precise but verbose. We can simplify it by
using shortcuts that rely on the fact that *every geom is associated
with a default statistic and position, and every statistic with a
default geom*. This means that you only need to specify one of stat or
geom to get a completely specified layer, with parameters passed on to
the geom or stat as appropriate. This expression generates the same
layer as the full layer command above:

```{r}
p <- ggplot(diamonds, aes(x = carat))
p <- p + geom_histogram(binwidth = 2, fill = "steelblue")
p
```


All the shortcut functions have the same basic form, beginning with
`geom\_` or `stat\_`:

```{r}
# arguments for geom_ or stat_
geom_XXX(mapping, data, ..., geom, position)
stat_XXX(mapping, data, ..., stat, position)
# mapping first, since almost always use the same data.
```


-   mapping (optional): A set of aesthetic mappings, specified using the
    `aes()` function and combined with the plot defaults.

-   data (optional): A dataset which overrides the default plot dataset.
    It is most commonly omitted, in which case the layer will use the
    default plot data.

-   ...: Parameters for the geom or stat, such as bin width in the
    histogram or bandwidth for a loess smoother. You can also use
    aesthetic properties as parameters. When you do this you set the
    property to a fixed value, not map it to a variable in the dataset.
    The example above showed setting the fill colour of the histogram to
    steelblue.

-   geom or stat (optional): You can override the default stat for a
    geom, or the default geom for a stat. This is a text string
    containing the name of the geom to use. Using the default will give
    you a standard plot; overriding the default allows you to achieve
    something more exotic

-   position (optional): Choose a method for adjusting overlapping
    objects.

Layers can be added to plots created with `ggplot()` or `qplot()`.
Remember, behind the scenes, `qplot()` is doing exactly the same thing:
it creates a plot object and then adds layers. The following example
shows the equivalence among these ways of making plots.

```{r}
qplot(sleep_rem/sleep_total, awake, data = msleep) + geom_smooth()
qplot(sleep_rem/sleep_total, awake, data = msleep, geom = c("point", "smooth"))
ggplot(msleep, aes(sleep_rem/sleep_total, awake)) + geom_point() + geom_smooth()
# store layer as variable, to reduce duplication
bestfit <- geom_smooth(method = "lm", se = F, colour = "steelblue", alpha = 1,
size = 1)
qplot(sleep_rem, sleep_total, data = msleep) + bestfit
qplot(awake, brainwt, data = msleep, log = "y") + bestfit
```


Data
----

The restriction on the data is simple: it must be a data frame. You can
replace the old dataset with `%+%`, as shown in the following example.

```{r}
p <- ggplot(mtcar, aes(mpg, wt, colour = cyl)) + geom_point()
p
mtcar <- transform(mtcars, mpg = mpg^2)
p %+% mtcar
```


Any change of values or dimensions is legitimate. However, if a variable
changes from discrete to continuous (or vice versa), you will need to
change the default scales, It is not necessary to specify a default
dataset except when using faceting; faceting is a global operation
(i.e., it works on all layers) and it needs to have a base dataset which
defines the set of facets for all datasets.

The data is stored in the plot object as a copy, not a reference. This
has two important consequences: if your data changes, the plot will not;
and ggplot2 objects are entirely self-contained so that they can be
save()d to disk and later load()ed and plotted without needing anything
else from that session.

Aesthetic mappings
------------------

We use the `ase` function to do this. You should never refer to
variables outside of the dataset (e.g., with diamonds\$carat). The
functions of variables can be
used, e.g. `aes(x=weight, y=height, colour=sqrt(age))`. The default
aesthetic mappings can be set when the plot is initialised or modified
later using +, as in this example:

```{r}
library(ggplot2)
# aesthetic mapping initially or later
p <- ggplot(mecars)
p <- p + aes(wt, hp)
p <- ggplot(mtcars, aes(x = mpg, y = wt))
p + geom_point()
```


The default mappings in the plot p can be extended or overridden in the
layers, as with the following code. Aesthetic mappings specified in a
layer affect only that layer. For that reason, unless you modify the
default scales, axis labels and legend titles will be based on the plot
defaults.

```{r}
# aesthetic mapping for layers only
p + geom_point(aes(colour = factor(cyl)))
p + geom_point(aes(y = disp))
## Setting VS. Mapping!
p <- ggplot(mtcars, aes(mpg, wt))
p + geom_point(colour = "darkblue") #setting
p + geom_point(aes(colour = "darkblue"))
# mapping, creat a new variable containing the value 'darkblue'.
qplot(mpg, wt, data = mtcars, colour = I("darkblue")) #use I() in qplot
```


Grouping
--------
```{r}
## Multiple groups, one aesthetic
library(nlme)
p <- ggplot(Oxboys, aes(age, height, group = Subject)) + geom_line()
## Different groups on different layers.
p + geom_smooth(aes(group = Subject), method = "lm", se = F)
# each subject has one smooth line
p + geom_smooth(aes(group = 1), method = "lm", se = F)
# group=1: use all data, one smooth line for all subjects.
```

```{r, eval=T,out.width='0.4 \ \textwidth',fig.show='hold'}
## Overriding the default grouping.
library(ggplot2)
library(nlme)
boysbox <- ggplot(Oxboys, aes(Occasion, height)) + geom_boxplot()
# Occasion is discrete
boysbox
boysbox + geom_line(aes(group = Subject), colour = "#3366FF")
# new group for new layer
```


Geoms
-----

Geometric objects, or geoms for short, perform the actual rendering of
the layer, control the type of plot that you create. For example, using
a point geom will create a scatterplot, while using a line geom will
create a line plot. Every geom has a default statistic, and every
statistic a default geom. For example, the bin statistic defaults to
using the bar geom to produce a histogram.

Stat
----

A statistical transformation, or stat, transforms the data, typically by
summarising it in some manner. For example, a useful stat is the
smoother, which calculates the mean of y, conditional on x, subject to
some restriction that ensures smoothness. A stat takes a dataset as
input and returns a dataset as output, and so a stat can add new
variables to the original dataset. It is possible to map aesthetics to
these new variables. For example, stat\_bin, the statistic used to make
histograms, produces the following variables: `count`, the number of
observations in each bin ; `density`, the density of observations in
each bin (percentage of total / bar width); `x`, the centre of the bin
These generated variables can be used instead of the variables present
in the original dataset. The names of generated variables must be
surrounded with` .. `when used. This prevents confusion in case the
original dataset includes a variable with the same name as a generated
variable.

```{r}
ggplot(diamonds, aes(carat)) + geom_histogram(binwidth = 0.1) #default use count as y axis
ggplot(diamonds, aes(carat)) + geom_histogram(aes(y = ..density..), binwidth = 0.1)
qplot(carat, ..density.., data = diamonds, geom = "histogram", binwidth = 0.1) #qplot
```


Pulling all together
--------------------

```{r}
## Combining geoms and stats
d <- ggplot(diamonds, aes(carat)) + xlim(0, 3)
d + stat_bin(aes(ymax = ..count..), binwidth = 0.1, geom = "area")
d + stat_bin(aes(size = ..density..), binwidth = 0.1, geom = "point", position = "identity")
model <- lme(height ~ age, data = Oxboys, random = ~1 + age | Subject) #linear mixed model
oplot <- ggplot(Oxboys, aes(age, height, group = Subject)) + geom_line()
age_grid <- seq(-1, 1, length = 10)
subjects <- unique(Oxboys$Subject)
preds <- expand.grid(age = age_grid, Subject = subjects)
# creat a data frame from all combinations of factors
Oxboys$fitted <- predict(model)
Oxboys$resid <- with(Oxboys, fitted - height)
oplot %+% Oxboys + aes(y = resid) + geom_smooth(aes(group = 1)) #apply to different data
```


Toolbox
=======

Basic plot types
----------------

```{r}
geom_area()
geom_bar(stat = "identity") #default multiple bars in same location stacked
geom_line()
geom_path() # in the order
geom_point() #scatterplot
geom_polygon()
geom_text() #need another aesthetic: label. hjust, vjust, angle
geom_tile()
# examples
df <- data.frame(x = c(3, 1, 5), y = c(2, 4, 6))
p <- ggplot(df, aes(x, y)) + xlab(NULL) + ylab(NULL)
p + geom_point() + ggtitle("geom_point")
9p + geom_bar(stat = "identity") + opts(title = "geom_bar(stat=\"identity\")")
p + geom_text(label = c("a", "b", "c")) + ggtitle("geom_text")
```


Displaying distributions
------------------------

```{r}
depthdist <- ggplot(diamonds, aes(depth)) + xlim(58, 68)
depthdist + geom_histogram(aes(y = ..density..), binwidth = 0.1) + facet_grid(cut ~.)
depthdist + geom_freqpoly(aes(y = ..density.., colour = cut), binwidth = 0.1)
```


Drawing maps
------------

```{r}
#maps
library(maps)
data(us.cities)
bigcities=subset(us.cities, pop>500000)
p=qplot(long,lat,data=bigcities) p+borders("state",size=.5)
wicities=subset(us.cities,country.etc=="WI")
ggplot(wicities,aes(long,lat))+borders("county","wisconsin",colour="grey70")+geom_point()
wi=map_data("county","wisconsin")
mid_range=function(x) mean(range(x,na.rm=TRUE))
library(plyr)
centres=ddply(wi, .(subregion),colwise(mid_range,.(lat,long)))
ggplot(wi, aes(long, lat)) + geom_polygon(aes(group = group),fill = NA, colour = "grey60") + geom_text(aes(label = subregion), data = centres, size = 4, angle = 0)
```


Toolbox
=======

Basic plot types
----------------

```{r, eval=TRUE,out.width='0.3 \ \textwidth',fig.show='hold'}
df <- data.frame(x = c(3, 1, 5), y = c(2, 4, 6), label = c("a", "b", "c"))
p <- ggplot(df, aes(x, y, label = label)) + xlab(NULL) + ylab(NULL)
p + geom_point() + ggtitle("geom_point")
p + geom_bar(stat = "identity") + ggtitle(("geom_bar(stat=\"identity\""))
p + geom_line() + ggtitle("geom_line")
p + geom_area() + ggtitle("geom_area")
p + geom_path() + ggtitle("geom_path")
p + geom_text() + ggtitle("geom_text")
p + geom_tile() + ggtitle("geom_tile")
p + geom_polygon() + ggtitle("geom_polygon")
```

```{r}
# xlim(low,high)
depth_dist <- ggplot(diamonds, aes(depth)) + xlim(58, 68)
depth_dist + geom_histogram(aes(y = ..density..), binwidth = 0.1) + facet_grid(cut ~
.)
depth_dist + geom_histogram(aes(fill = cut), binwidth = 0.1, position = "fill")
depth_dist + geom_freqpoly(aes(y = ..density.., colour = cut), binwidth = 0.1)
qplot(cut, depth, data = diamonds, geom = "boxplot")
qplot(carat, depth, data = diamonds, geom = "boxplot", group = round_any(carat,
0.1, floor), xlim = c(0, 3))
```

Dealing with overplotting
-------------------------

```{r}
# shape or fuzzing
df <- data.frame(x = rnorm(2000), y = rnorm(2000))
norm <- ggplot(df, aes(x, y))
norm + geom_point()
norm + geom_point(shape = 1) # hollow points
norm + geom_point(shape = ".") # Pixel sized
norm + geom_point(colour = "black", alpha = 1/5)
# Jitter
td <- ggplot(diamonds, aes(table, depth)) + xlim(50, 70) + ylim(50, 70)
td + geom_point()
td + geom_jitter()
jit <- position_jitter(width = 0.5)
td + geom_jitter(position = jit)
td + geom_jitter(position = jit, colour = "black", alpha = 1/5)
```


Scales, axes and legends
========================

Scales: common arguments
------------------------

```{r}
p <- qplot(cty, hwy, data = mpg, colour = displ) 
p + scale_x_continuous("City mpg") 
p + xlab("City mpg") 
p + ylab("Highway mpg") 
p + labs(x = "City mpg", y = "Highway", colour = "Displacement") 
p + xlab(expression(frac(miles, gallon)))
```


Breaks controls which values appear on the axis or legend, i.e., what
values tick marks should appear on an axis or how a continuous scale is
segmented in a legend. labels specifies the labels that should appear at
the breakpoints. If labels is set, you must also specify breaks, so that
the two can be matched up correctly. To distinguish breaks from limits,
remember that breaks affect what appears on the axes and legends, while
limits affect what appears on the plot.

```{r}
p <- qplot(cyl, wt, data = mtcars) 
p + scale_x_continuous(breaks = c(5.5, 6.5)) 
p + scale_x_continuous(limits = c(5.5, 6.5)) 
p <- qplot(wt, cyl, data = mtcars, colour = cyl) 
p + scale_colour_gradient(breaks = c(5.5, 6.5)) 
p + scale_colour_gradient(limits = c(5.5, 6.5))
```


Facet
-----

There are two types of faceting provided by ggplot2: `facet_grid` and
`facet_wrap`. Facet grid produces a 2d grid of panels defined by
variables which form the rows and columns, while facet wrap produces a
1d ribbon of panels that is wrapped into 2d.

You can access either faceting system from qplot(). A 2d faceting
specification (e.g., x \~ y) will use `facet_grid`, while a 1d
specification (e.g., \~ x) will use `facet_wrap`. For `facet_grid` there
is an additional parameter called space, which takes values "free"
or "fixed". When the space can vary freely, each column (or row)
will have width (or height) proportional to the range of the scale for
that column (or row). This makes the scaling equal across the whole
plot.`facet_grid(manufacturer ~ ., scales = `"free`", space = `"free`")`

```{r}
mpg2 <- subset(mpg, cyl != 5 & drv %in% c("4", "f")) 
qplot(cty, hwy, data = mpg2) + facet_grid(. ~ cyl) #a single row with multiple column 
qplot(cty, data = mpg2, geom="histogram") + facet_grid(cyl ~ .) #a single col with multiple row 
#facet_grid(variable ~ ., scales = "free_y")
qplot(cty, hwy, data = mpg2) + facet_grid(drv ~ cyl) #multiple rows and cols 
p <- qplot(displ, hwy, data = mpg2) +  geom_smooth(method = "lm", se = F) 
p + facet_grid(cyl ~ drv) 
p + facet_grid(cyl ~ drv, margins = T) #Margin like contingency table
qplot(displ, hwy, data = mpg2) + geom_smooth(aes(colour = drv), method = "lm", se = F) + facet_grid(cyl ~ drv, margins = T)
```


### Group VS facet

```{r, eval=TRUE,out.width='0.4 \ \textwidth',fig.show='hold'}
dplot <- ggplot(diamonds, aes(carat, price, colour = color)) 
dplot+geom_point() 
dplot + geom_point() + facet_grid(. ~ color) 
```

### Dodging vs. faceting

```{r,eval=TRUE,out.width='0.4 \ \textwidth',fig.show='hold'}
qplot(color, data=diamonds, geom = "bar", fill = cut, position="dodge") 
qplot(cut, data = diamonds, geom = "bar", fill = cut) + facet_grid(. ~ color) +theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 8,colour = "grey50")) 
```


Legend
------

```{r}
##change default point color for discrete variables.
cbPalette <- c("blue", "red") 
p+geom_point(data=data,aes(long,lat,colour=type,shape=type))+theme_bw()+theme(legend.position="top")+scale_colour_manual(values=cbPalette) 
```


Theme
-----

```{r}
ggplot()+theme_bw()
```


plyr package
------------

```{r,eval=TRUE}
library(plyr) 
head(ddply(diamonds, .(color), subset, carat == min(carat))) #subset to select data 
head(ddply(diamonds, .(color), transform, price = scale(price))) #transform to change 
head(ddply(diamonds, .(color), transform, price = price - mean(price))) 
#colwise() returns a new function 
nmissing <- function(x) sum(is.na(x)) 
nmissing(msleep$brainwt) 
nmissing_df <- colwise(nmissing) 
nmissing_df(msleep) 
colwise(nmissing)(msleep) # a shortcut 
#numcolwise(). catcolwise() only works with categorical columns 
ddply(msleep, .(vore), numcolwise(median), na.rm = T) 
my_summary <- function(df) {with(df, data.frame(pc_cor = cor(price, carat, method = "spearman"), lpc_cor = cor(log(price), log(carat))))} 
ddply(diamonds, .(cut), my_summary)
```


